---
title: 'Play with the tribe of attributes'
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Play with the tribe of attributes}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(dplyr)
library(tribe)
print.list <- function(x, ...) str(x)
```


## Verbs for easy manipulation of attributes

The tribe package provides verbs for easy manipulation of attributes. 
These verbs are: 

  - `at_mutate` to create or modify attributes; 
  - `at_select` to select attributes (and `NULL`ify the others);
  - `at_rename` to rename attributes; 
  - `at_slice` to extract attributes. 

The function `tribe` is a convenient synonym of `attributes`, 
with the slight difference that it always returns a named list. 

```{r}
df <- data.frame(x = 1:2, y = 2:3) %>%
  at_mutate(example="yes", package="dplyr")
tribe(df)
```

Use `at_slice` to extract attribute values: 

```{r}
at_slice(df, names)
```

Each verb has its standard evaluation version; for instance `at_slice_` is the standard evaluation version of `at_slice`: 

```{r}
at_slice_(df, "class")
at_slice_(df, ~ package)
```

Similarly `at_mutate_` is the standard evaluation version of `at_mutate`: 

```{r}
df <- df %>%
  at_mutate_(package = ~ NULL, # deletes the attribute called 'package'
             example = ~ "no")
tribe(df)
```


## A new pipe that preserves attributes

The tribe package builds on the magrittr package and brings a new pipe `%@>%` similar to the pipe `%>%` that enables propagation of attributes. 

```{r}
df <- data.frame(x = 1:2, y = 2:3) %>%
 at_mutate(example="yes",
           package="tribe", 
           class = c("my_tbl", "data.frame"))
```

Attributes just created are often lost when the object passes through dplyr verbs, for instance: 

```{r}
tribe(df %>% mutate(z=3))
```

With the new pipe `%@>%`, most attributes propagate: 

```{r}
tribe(df %@>% mutate(z=3))
```

Behind the scene, the `shield` function operates: 

```{r}
# Attributes are lost when the object passes through dplyr verbs
df2 <- df %>% mutate(z = 3)
tribe(df2)

# Most attributes are kept
df3 <- shield(df2, tribe(df), propagate = "most")
tribe(df3)

# To keep the class, use 'keep_also'
df4 <- shield(df2, tribe(df), propagate = "most", keep_also = "class")
tribe(df4)
```

One can create a new pipe to adjust attributes propagation settings: 

```{r, warning = FALSE}
"%newpipe>%" <- make_pipe(propagate="none", keep_also = "example")
tribe(df %newpipe>% mutate(z=3))
```
